NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-02-08 Manuscript/manuscriptPS.tex
Update manuscriptPS.tex

\begin{figure}[!tb]
  \centering
  \includegraphics[width=8.5cm]{../Figs/fig1_POPS.pdf}
  \caption{\label{PShgSIGNSsimpson}
    The headgroup and glycerol backbone region of the (A) INEPT spectrum and
    (B) 2D R-PDLF spectra.
    (C) Experimental S-DROSS data (points), and SIMPSON simulations (blue lines) with
    the order parameter values of -0.12 for the $\beta$-carbon, and 0.09 and -0.02
    for the $\alpha$-carbon splittings.
    Dashed gray line is the S-DROSS curve from a SIMPSON simulation with a positive value (+0.02) 
    for the smaller $\alpha$-carbon order parameter.
  }
  \todo{I think that the peak labeling would be good to show also in (A).}\\
  \todo{Also for $\alpha$ these are OPs, right,  thus we could just exclude the word 'splittings'?}
  \todo{Please confirm: Was the positive value used for $\alpha$ in the gray curve 0.02?}
\end{figure}
\begin{figure}[!htb]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PCPScomp.eps}
  \caption{\label{HGorderParameters}
    (A) Chemical structures and labels for the headgroup and glycerol backbone carbons.
    (B) Headgroup and glycerol backbone order parameters of POPS (T = 298 K) measured in this work compared
    with the previously published values from DOPS (T = 303 K, $^2$H NMR, 0.1M of NaCl) \cite{browning80} and 
    POPC  (T = 300 K, $^{13}$C NMR) \cite{ferreira13} experiments. Signs of the PS order parameters
    are measured in this work. Signs of the PC order parameters are measured previously~\cite{ferreira16}.
  }\todo{
  1) Use diamonds for DOPS, spheres for POPS.
  2) Error bars?
  3) Change the y-label to $S_\mathrm{CH}$, and invert the y-axis, as in Fig. 3.
  4) Make lower g1 POPC visible, e.g., by slightly larger point.
  }
\end{figure}

@@

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPS.pdf}
  \caption{\label{HGorderParametersPS}
    Order parameters of PS headgroup ($\beta$ and $\alpha$) and
    glycerol backbone (g$_3$, g$_2$, g$_1$) from NMR experiments (horizontal lines),
    and MD simulations with different force fields (dots).
    Experimental data for DOPS are measured with 0.1~M of NaCl~\cite{browning80},
    while all the other data are without additional salt.
    All the data for DOPS at 303~K and all the data for POPS at 298~K.
    Light blue areas span 0.04 units around the average of the extremal experimental values,
    in accordance with the expected quantitative accuracy of experimental values~\cite{ollila16}.
    The vertical bars shown for all simulation values except MacRog~K$^+$
    are not error bars, but demonstrate that for these systems
    we had at least two data sets; the ends of the bars mark the extreme values
    from the sets, and the dot marks their measurement-time-weighted average.
  }
\end{figure}

The headgroup and glycerol backbone order parameters of 
POPS measured in this work are in good agreement with the previously reported
values from $^2$H\,NMR experiments of DOPS \cite{browning80} (Fig. \ref{HGorderParameters}).
When compared with the previously measured values for POPC \cite{ferreira13} (Fig. \ref{HGorderParameters}),
the $\beta$-carbon order parameter is significantly more negative and $\alpha$-carbon
experiences a significant forking (different order parameters for the two hydrogens in the same carbon \cite{ollila16}) in the PS headgroup.
These features have been intepreted to arise from a rigid PS headgroup
conformation, stabilized by hydrogen bonds or electrostatic
interactions \cite{browning80,buldt81}, but detailed structrural interpretation is not
available.

>>>>

\begin{figure}[!tb]
  \centering
  \includegraphics[width=8.5cm]{../Figs/fig1_POPS.pdf}
  \caption{\label{PShgSIGNSsimpson}
    The headgroup and glycerol backbone region of the (A) INEPT spectrum and
    (B) 2D R-PDLF spectra.
    (C) Experimental S-DROSS data (points), and SIMPSON simulations (blue lines) with
    the order parameter values of -0.12 for the $\beta$-carbon, and 0.09 and -0.02
    for the $\alpha$-carbon splittings.
    Dashed gray line is the S-DROSS curve from a SIMPSON simulation with a positive value (+0.02) 
    for the smaller $\alpha$-carbon order parameter.
  }
  \todo{I think that the peak labeling would be good to show also in (A).}\\
  \todo{Also for $\alpha$ these are OPs, right,  thus we could just exclude the word 'splittings'?}
  \todo{Please confirm: Was the positive value used for $\alpha$ in the gray curve 0.02?}
\end{figure}
\begin{figure}[!htb]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PCPScomp.eps}
  \caption{\label{HGorderParameters}
    (A) Chemical structures and labels for the headgroup and glycerol backbone carbons.
    (B) Headgroup and glycerol backbone order parameters of POPS (T = 298 K) measured in this work compared
    with the previously published values from DOPS (T = 303 K, $^2$H NMR, 0.1M of NaCl) \cite{browning80} and 
    POPC  (T = 300 K, $^{13}$C NMR) \cite{ferreira13} experiments. Signs of the PS order parameters
    are measured in this work whereas signs of the PC order parameters are measured previously~\cite{ferreira16}.
  }\todo{
  1) Use diamonds for DOPS, spheres for POPS.
  2) Error bars?
  3) Change the y-label to $S_\mathrm{CH}$, and invert the y-axis, as in Fig. 3.
  4) Make lower g1 POPC visible, e.g., by slightly larger point.
  }
\end{figure}

@@

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPS.pdf}
  \caption{\label{HGorderParametersPS}
    Order parameters of PS headgroup ($\beta$ and $\alpha$) and
    glycerol backbone (g$_3$, g$_2$, g$_1$) from NMR experiments (horizontal lines),
    and MD simulations with different force fields (symbols).
    Experimental data for DOPS are measured with 0.1~M of NaCl~\cite{browning80},
    while all the other data are without additional salt.
    The data for DOPS is at 303~K and the data for POPS is at 298~K.
    Light blue areas span 0.04 units around the average of the extremal experimental values,
    in accordance with the expected quantitative accuracy of experiments~\cite{ollila16}.
    The vertical bars shown for all simulation values (excl. MacRog~K$^+$)
    are not error bars, but demonstrate that for these systems
    we had at least two data sets; the ends of the bars mark the extreme values
    from the sets, and the dot marks their measurement-time-weighted average.
  }
\end{figure}

The headgroup and glycerol backbone order parameters of 
POPS measured in this work are in good agreement with the previously reported
values from $^2$H\,NMR experiments of DOPS \cite{browning80} (Fig. \ref{HGorderParameters}).
When compared with the previously measured values for POPC \cite{ferreira13} (Fig. \ref{HGorderParameters}),
the $\beta$-carbon order parameter is significantly more negative and $\alpha$-carbon
experiences a significant forking (different order parameters for the two hydrogens in the same carbon \cite{ollila16}) in the PS headgroup.
These features have been intepreted to arise from a rigid PS headgroup
conformation, stabilized by hydrogen bonds or electrostatic
interactions \cite{browning80,buldt81}, but detailed structrural interpretation is not
available. 
%Moved here from the molecular electrometer paragraph because there it distrupted the flow of the section and it contains results. This also enables referring to figures in order. I also removed the refs which do not seem to refer to the actual data used in the figure.
%Rewrote the paragraph, see if ok.

We note that the the DOPS $^2$H~NMR reference data found in the literature~\cite{browning80,roux90} was obtained by first solvating the lipids to the buffer and then centrifuging the sample to a pellet that was used in the measurements. Such samples have a lower lipid concentration
(approximately 10~wt~\% of lipids~\cite{browning80,roux88,roux90}) than 
gravimetric samples (60~wt~\%) and simulations (approximately 50-60~wt~\%) in this work.
Larger multilamellar repeat distances are expected in the samples with lower lipid
concentrations due to the swelling caused by electrostatic repulsion in pure PS lipid systems~\cite{millman82}.
Yet the PS headgroup order parameters measured from gravimetric samples (POPS) in this work
are in good agreement with the results from centrifuged samples~\cite{browning80}. This, together with the rapid decrease of equilibrium repeat distance with addition of monovalent salt~\cite{millman82,rand89}, indicates that the hydration levels of multilamellae are sufficiently similar in the simulations and reference experiments.
